Skew-symmetric entropy stable modal discontinuous Galerkin formulations

### 2019-09-06 docs/skewform/main.tex
updating docs

\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

@@

\makeatletter
\renewcommand*{\@thmcounterend}{}
\makeatother
%\spnewtheorem{lemma}{Lemma}{\bfseries}{\itshape}
\spnewtheorem{assumption}{Assumption}{\bfseries}{\itshape}
%\spnewtheorem{remark}{Remark}{\bfseries}{\itshape}
\usepackage{cite}

\usepackage{slashbox} % for slashbox in tables

\usepackage{amsmath, amssymb, amsfonts}
%\newtheorem*{remark}{Remark}
%\theoremstyle{assumption}
%\newtheorem{assumption}{Assumption}
\usepackage{hyperref}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{tikz}
\usepackage{hhline}
\usepackage{graphicx, color, bm}
\usepackage{subfig}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{slopetri} % my own package for drawing triangle slopes

@@

\renewcommand{\note}[1]{{\color{blue}{#1}}}

@@

\begin{abstract}
Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.  In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which a traditional SBP property does not hold.  These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.   We conclude with an application of the skew-symmetric formulation for entropy stable schemes on mixed quadrilateral-triangle meshes. 
\end{abstract}

\tableofcontents

@@

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.   By using an approach referred to as ``flux differencing'', entropy stable schemes account for the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.  Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.

Entropy stable discretizations are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.  However, the SBP property does not hold for some under-integrated quadrature rules, which naturally arise in certain discretization settings.  For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.  On degree $N$ tensor product elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.  When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.  However, while GLL quadrature induces an SBP property on quadrilateral elements, but does not always result in an SBP property if used for triangular elements \cite{chan2017discretely}.  %A similar issue arises on non-conforming meshes, where a composite ``split-side'' quadrature naturally when coupling larger elements to smaller elements through a non-conforming interface \cite{kozdon2018energy}.  For quadrilaterals utilizing a GLL volume quadrature, the use of a composite quadrature on a non-conforming face also results in the loss of the SBP property \cite{chan2017discretely}.

This work proposes an alternative formulation which utilizes a skew-symmetric form of flux differencing.  Under such a formulation, the proof of entropy stability requires only a weaker ``variational'' form of the SBP property, which holds under a more general class of quadrature rules compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.  We show that the skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.

It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.  These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature and can be computed either algebraically or through an optimization problem.  The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that we start with an explicit approximation space, which then induce quadrature-based operators.  The formulations presented here also accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.

The structure of the paper is as follows: Section~\ref{sec:nonlin} describes the continuous entropy inequality which we aim to replicate discretely.  Section~\ref{sec:approx} and Section~\ref{sec:sbp} introduce polynomial approximation spaces and quadrature-based SBP operators on simplicial and tensor product elements.  Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.  Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:num} presents numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.

@@

\section{Quadrature-based matrices and decoupled SBP operators}
\label{sec:sbp}
%\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.  These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.  We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.

@@

Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.  Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{gather}
\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, \label{eq:QN}\\
\bm{E} = \bm{V}_f\bm{P}_q, \qquad \bm{B}^i = \bm{W}_f \diag{\hat{\bm{n}}_i}, \qquad \bm{Q}^i_N  = \LRs{
\begin{array}{cc}
\bm{Q}^i - \frac{1}{2}\bm{E}^T\bm{B}^i\bm{E} &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
-\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
\end{array}}.  \nonumber
\end{gather}
Here, $\bm{B}^i$ is a boundary ``integration'' matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.

It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
%The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$, 
%\begin{align*}
%\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
%\end{align*}

For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
\begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
Assume that $\bm{Q}^i$ satisfies a generalized SBP property
\[
\bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T.
\]
Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
\begin{gather}
\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{B}^i\end{array}}.\label{eq:dsbp}
\end{gather}
%\begin{enumerate}
%\item The mass matrix is positive definite, 
%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
%\end{enumerate}
\label{lemma:dsbp}
\end{theorem}
\begin{proof}
The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
\end{proof}

@@

\subsection{A variational SBP property}

@@

The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.  We will relax accuracy conditions on these quadrature rules by utilizing a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.

@@

\begin{lemma}{(Discrete FTC and exact differentiation of constants)}
\label{lemma:sbpcor}
Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
\[
\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0}.
\]
\end{lemma}
\begin{proof}
The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.     The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} 
\[
\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}^i_N}\bm{u}.
\]
\end{proof}

@@

\end{proof}

@@

The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is sufficient to show the numerical solution convergences to the weak solution under mesh refinement.  
\begin{theorem}
The formulation (\ref{eq:esdgSkew}) is locally conservative such that
\begin{align}
\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\bm{f}_i^* = 0. 
\end{align}
\end{theorem}
\begin{proof}
To show local conservation, we test (\ref{eq:esdgSkew}) with $1$
\begin{align}
\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}_h}{t} + \sum_{i=1}^d
\bm{1}^T
\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
\end{align}
Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.  Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term $\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1}$ vanishes.
% and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
%\begin{align}
%\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - \bm{1}^T\LRp{\LRp{\bm{Q}^i_N}^T\circ \bm{F}_S}\bm{1}\\
%&= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - 
%\bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} = 0,
%\end{align}
%where we have used that $\bm{F}_S$ is symmetric.
%\begin{align}
%\end{align}
\end{proof}

@@

\section{Conditions on the accuracy of volume and surface quadrature rules}

The previous sections outline minimal conditions under which entropy stability is guaranteed under a skew-symmetric formulation and a polynomial geometric mapping.  In Section~\ref{sec:curvgeo}, we translate these minimial conditions into conditions on quadrature accuracy.

@@

\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
\label{sec:curvgeo}

@@

A low storage 4th order Runge-Kutta scheme \cite{carpenter1994fourth} is used for all numerical experiments.  The time-step is estimated based on formulas derived for linear advection in \cite{chan2015gpu, chan2018multi}
\[
dt = C \frac{h}{c_{\max} \max\LRc{\frac{1}{2}C_T, C_I}}
\]
where $h$ is the mesh size, $c_{\max}$ is the maximum wave-speed, $C$ is a user-defined CFL constant, and $C_T, C_I$ are constants in finite element inverse and trace inequalities.  These constants scale proportionally to $N^2$, though precise values of $C_I, C_T$ vary slightly depending on the choice of volume or surface quadrature used.  The dependence of $C_I, C_T$ on quadrature is discussed in more detail in Appendix~\ref{sec:consts}, where computed values of $C_I, C_T$ are also given.

>>>>

\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%\usepackage[T1]{fontenc}

@@

\makeatletter
\renewcommand*{\@thmcounterend}{}
\makeatother
%\spnewtheorem{lemma}{Lemma}{\bfseries}{\itshape}
\spnewtheorem{assumption}{Assumption}{\bfseries}{\itshape}
%\spnewtheorem{remark}{Remark}{\bfseries}{\itshape}

%\usepackage{fullpage}
\usepackage{cite}

%\usepackage{slashbox} % for slashbox in tables

\usepackage{amsmath, amsfonts, amssymb}
%\newtheorem*{remark}{Remark}
%\theoremstyle{assumption}
%\newtheorem{assumption}{Assumption}
\usepackage{hyperref}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{tikz}
%\usepackage{hhline}
\usepackage{graphicx, color, bm}
\usepackage{subfig}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{slopetri} % my own package for drawing triangle slopes

@@

\renewcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\bnote}[1]{{\color{blue}{#1}}}
\newcommand{\rnote}[1]{{\color{red}{#1}}}

@@

\begin{abstract}
High order entropy stable discontinuous Galerkin (DG) methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  The construction of such schemes has relied on the use of carefully chosen nodal points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.  In this work, we show how to construct ``modal'' DG formulations which are \bnote{entropy stable for volume and surface quadratures under which the SBP property in \cite{chan2017discretely} does not hold.  These formulations rely on an alternative skew-symmetric construction of operators which automatically satisfy the SBP property.   Entropy stability then follows for choices of volume and surface quadrature which satisfy sufficient accuracy conditions.}   \rnote{The accuracy of these new SBP operators is shown to depend on a separate set of accuracy conditions. }
%avoid the use of a ``strong'' matrix-based SBP property, and instead rely on the assumption that discrete operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.   
We conclude with numerical experiments verifying the accuracy and stability of the proposed formulations, and discuss an application of these formulations for entropy stable DG schemes on mixed quadrilateral-triangle meshes.  
\end{abstract}

%\tableofcontents

@@

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.   Entropy stable schemes circumvent the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality by using an approach referred to as ``flux differencing''.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.  Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.

Entropy stable ``modal'' DG discretizations \cite{chan2017discretely, chan2018discretely} are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.  However, the SBP property does not hold for certain under-integrated quadrature rules, which arise naturally in some discretization settings.  For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.  On degree $N$ quadrilateral elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.  When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.  However, while GLL quadrature induces an SBP property on quadrilateral elements, it does not guarantee an SBP property if used on triangular elements \cite{chan2017discretely}.

This work proposes an alternative formulation which utilizes a \note{skew-symmetric construction of the SBP operator}.  Under such a formulation, the proof of entropy stability \note{holds under weaker quadrature rules} compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.  We show that this skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.

It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which satisfy similar accuracy conditions and an SBP property \cite{hicken2016multidimensional}.  These operators exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature, but do not correspond to any specific basis or approximation space.  The formulations in \cite{chen2017entropy, crean2018entropy} differ from the ones presented in this work in that they are based on SBP finite differences and ``nodal'' (rather than ``modal'') DG formulations, with differentiation operators computed algebraically or through an optimization problem for each specific choice of nodes.  In contrast, ``modal'' formulations induce quadrature-based operators from an explicit approximation space, and accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated quadrature rules).

The structure of the paper is as follows: Section~\ref{sec:nonlin} describes the continuous entropy inequality which we aim to replicate discretely.  Section~\ref{sec:approx} and Section~\ref{sec:sbp} introduce polynomial approximation spaces and quadrature-based SBP operators on simplicial and tensor product elements.  
Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.  Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:quadacc} provides explicit conditions for entropy stability in terms of quadrature accuracy and the polynomial degree of geometric mappings.  Section~\ref{sec:num} concludes by presenting numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.

@@

\section{Quadrature-based matrices and ``hybridized'' SBP operators}
\label{sec:sbp}
%\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.  These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.  We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.

@@

Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.  Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$.  The ``hybridized'' (originally referred to as ``decoupled'' in \cite{chan2017discretely}) summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{gather}
\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, \label{eq:QN}\\
\bm{E} = \bm{V}_f\bm{P}_q, \qquad \bm{B}^i = \bm{W}_f \diag{\hat{\bm{n}}_i}, \qquad \bm{Q}^i_N  = \LRs{
\begin{array}{cc}
\bm{Q}^i - \frac{1}{2}\bm{E}^T\bm{B}^i\bm{E} &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
-\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
\end{array}}.  \nonumber
\end{gather}
Here, $\bm{B}^i$ is a boundary ``integration'' matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.

It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then, $f\pd{g}{x_i}$ can be approximated by the polynomial with modal coefficients given by
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
%The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$, 
%\begin{align*}
%\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
%\end{align*}

For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
\begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
If $\bm{Q}^i$ satisfies a generalized SBP property
\[
\bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T,
\]
then the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
\begin{gather}
\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{B}^i\end{array}}.\label{eq:dsbp}
\end{gather}
%\begin{enumerate}
%\item The mass matrix is positive definite, 
%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
%\end{enumerate}
\label{lemma:dsbp}
\end{theorem}
\begin{proof}
The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
\qed\end{proof}

@@

\subsection{\bnote{An alternative construction of SBP operators}} % variational SBP property}

%\subsection{A variational SBP property}

@@

The property (\ref{eq:dsbp}) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.  We will relax accuracy conditions on these quadrature rules by utilizing \note{an alternative construction of $\bm{Q}^i_N$ based the skew-symmetric matrix $\bm{Q}^i - \LRp{\bm{Q}^i}^T$:}
%a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.  
\bnote{\begin{equation}
\tilde{\bm{Q}}^i_{N} = \begin{bmatrix}
\bm{Q}^i - \LRp{\bm{Q}^i}^T &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
-\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
\end{bmatrix}.
\label{eq:skewQN} 
\end{equation}
It is clear to see that the skew-symmetric construction of $\tilde{\bm{Q}}^i_N$ as defined in (\ref{eq:skewQN}) satisfies the SBP property (\ref{eq:dsbp}) by construction.  However, the accuracy of $\tilde{\bm{Q}}^i_N$ as a differentiation operator becomes slightly more difficult to analyze. }

@@

\begin{lemma}{(Discrete FTC and exact differentiation of constants)}
\label{lemma:sbpcor}
Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
\[
\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0}.
\]
\end{lemma}
\begin{proof}
The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.     The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} 
\[
\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}^i_N}\bm{u}.
\]
\qed\end{proof}

@@

\qed\end{proof}

@@

The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is sufficient to show the numerical solution convergences to the weak solution under mesh refinement.  
\begin{theorem}
The formulation (\ref{eq:esdgSkew}) is locally conservative such that
\begin{align}
\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\bm{f}_i^* = 0. 
\end{align}
\end{theorem}
\begin{proof}
To show local conservation, we test (\ref{eq:esdgSkew}) with $1$
\begin{align}
\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}_h}{t} + \sum_{i=1}^d
\bm{1}^T
\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
\end{align}
Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.  Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term $\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1}$ vanishes.
% and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
%\begin{align}
%\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - \bm{1}^T\LRp{\LRp{\bm{Q}^i_N}^T\circ \bm{F}_S}\bm{1}\\
%&= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - 
%\bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} = 0,
%\end{align}
%where we have used that $\bm{F}_S$ is symmetric.
%\begin{align}
%\end{align}
\qed\end{proof}

@@

\section{Conditions on the accuracy of volume and surface quadrature rules}
\label{sec:quadacc}

The previous sections outline minimal conditions under which entropy stability is guaranteed under a skew-symmetric formulation and a polynomial geometric mapping.  In Section~\ref{sec:quadacc}, we translate these minimial conditions into conditions on quadrature accuracy.

@@

\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}

@@

A low storage 4th order Runge-Kutta scheme \cite{carpenter1994fourth} is used for all numerical experiments.  The time-step is estimated based on formulas derived for linear advection in \cite{chan2015gpu, chan2018multi}
\[
dt = C \frac{h}{c_{\max} \max\LRc{\frac{1}{2}C_T, C_I}}
\]
where $h$ is the mesh size, $c_{\max}$ is the maximum wave-speed, $C$ is a user-defined CFL constant, and $C_T, C_I$ are constants in finite element inverse and trace inequalities.  These constants scale proportionally to $N^2$, though precise values of $C_I, C_T$ vary slightly depending on the choice of volume or surface quadrature used.  The dependence of $C_I, C_T$ on quadrature is discussed in more detail in Appendix~\ref{sec:consts}, where computed values are also given.
